---
title: "class08"
author: "tiantaima"
date: "10/25/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## k means example
```{r}
 # Generate some example data for clustering
tmp <- c(rnorm(30,-3), rnorm(30,3)) 
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```


 ##Use the kmeans() function setting k to 2 and nstart=20 Inspect/print the results
Q. How many points are in each cluster?
Q. What ‘component’ of your result object details
- cluster size?
- cluster assignment/membership? - cluster center?
Plot x colored by the kmeans cluster assignment and
      add cluster centers as blue points
```{r}
k <- kmeans(x,centers = 2, nstart = 20)
k
table(k$cluster)
plot(x, col=k$cluster)
points(k$centers, col="blue", pch=15)
```

## Hierarchical clustering in R

```{r}
hc <- hclust(dist(x))

```

## plot my result
```{r}
plot(hc)

abline(h=4, col="blue")
cutree(hc, h=4)

```
```{r}
greps <- cutree (hc, h=4)
table(greps)
```

```{r}
greps <- cutree(hc,k=2)
greps
```
```{r}
plot(x, col = greps)
```





```{r}
 #Step 1. Generate some example data for clustering
x <- rbind(
matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2), # c1
matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
matrix(c(rnorm(50, mean = 1, sd = 0.3), # c3
rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x) <- c("x", "y")

# Step 2. Plot the data without clustering
plot(x)

# Step 3. Generate colors for known clusters
# (just so we can compare to hclust results) 
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x, col=col)


# cluster 
hc <- hclust(dist(x))
plot(hc)
abline(h=2.9, col="blue")
greps <- cutree(hc, k=3)
plot(x, col=greps)
table(col,greps)
```

```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV", row.names=1)
mydata

pca <- prcomp(t(mydata), scale=TRUE)
attributes(pca)
 
```
```{r}
pca$x[,1]
plot(pca$x[,1],pca$x[,2])
```
```{r}
summary(pca)
```

```{r}
## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
pca.var.per
```

```{r}
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
barplot(pca.var.per, main="Scree Plot",
xlab="Principal Component", ylab="Percent Variation")
```

```{r}
# Alternatively 
plot(pca)
```


```{r}
plot(pca$x[,1:2], col=c("red","red","red","red","red","blue","blue","blue","blue","blue"))

```

```{r}
rep("red",5)
```
```{r}

```

```{r}
x <- read.csv("UK_foods.csv")
x
nrow(x)
ncol
```
```{r}
head(x)
tail(x)
```

```{r}
rownames(x) <- x[,1]
x <- x[,-1]
head(x)
dim(x)
```

```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```

```{r}
pairs(x, col=rainbow(10), pch=16)
```

```{r}
pca <- prcomp( t(x) )
summary(pca)
pca
```

```{r}
# Plot PC1 vs PC2
plot(pca$x[,1], pca$x[,2], xlab="PC1", ylab="PC2", xlim=c(-270,500),pch=1, cex=0.5)
text(pca$x[,1], pca$x[,2], colnames(x), col=c(1,2,3,4))
```

```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
```

```{r}
z <- summary(pca)
z$importance
```

```{r}
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```

```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```

```{r}
## The inbuilt biplot() can be useful for small datasets 
biplot(pca)         
```

