---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup Bioconductor DESeq2

```{r eval=FALSE}
install.packages("BiocManager")
BiocManager::install()
BiocManager::install("DESeq2")
```

## Import data for today's class 

```{r}
counts <- read.csv("airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <- read.csv("airway_metadata.csv", stringsAsFactors = F)
```


## Have a look 

```{r}
head(counts)
```

## How many genes are in this dataset? 

```{r}
nrow(counts)
```

## How many experiments? 

```{r}
ncol(counts)-1
```


Let make sure metadata id col match the colnames of the counts 

```{r}
colnames(counts)[-1]
metadata$id
```

```{r}
colnames(counts)[-1]==metadata$id
# or use all(), the all()function is useful for looking for an arrary of logicals 
all(colnames(counts)[-1]==metadata$id)
```

## Analysis : compare the control to the drug treated

First we need to find the columns of our couontData athat are control and then finf their mean 

```{r}
metadata$dex=="control"
```

```{r}
control.id <-metadata[metadata$dex=="control",]$id

```

focus on the counts for control experiment: 
```{r}
head(counts[,control.id])
```

Mean counts for control experiemnt

```{r}
control.mean <- rowSums(counts[,control.id])/length(control.id)
names(control.mean) <- counts$ensgene
```





Then we analyze the treated columns: 

```{r}
treated.id <- metadata[metadata$dex=="treated", ]$id
```

```{r}
head(counts[,treated.id])
```



```{r}
treated.mean <- rowSums(counts[,treated.id])/length(treated.id)
names(treated.mean) <- counts$ensgene
```


Lets store them together: 

```{r}
meancounts <- data.frame(control.mean, treated.mean)
```

Plot control VS. treated: 

```{r}
plot(meancounts$control.mean, meancounts$treated.mean)
```

Good case of a log plot:

```{r}
plot(meancounts$control.mean, log2(meancounts$treated.mean))
```



## Fold change
```{r}
meancounts$log2fc <-log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

There are a couple of “weird” results. Namely, the NaN ("not a number") and -Inf (negative infinity) results.

The NaN is returned when you divide by zero and try to take the log. The -Inf is returned when you try to take the log of zero. It turns out that there are a lot of genes with zero expression. Let’s filter our data to remove these genes. Again inspect your result (and the intermediate steps) to see if things make sense to you

```{r}
# Test for finding 0 entries
x <- data.frame(happy=c(5,6,0,0), sad=c(0,5,5,0))
x
x==0
```

```{r}
which(x==0)
which(x==0, arr.ind = T)
```

Focus on the rows only have zero entries: 
```{r}
dd<-unique(which(x==0, arr.ind = T))
dd      
inds <- unique(which(x==0, arr.ind = T)[,1])
inds
x[-inds,]
```

In the gene dataset: 

```{r}
to.rm <- unique(which(meancounts[,1:2]==0, arr.ind=T)[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```




```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

```{r}
sum(up.ind)
sum(down.ind)
```



```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```
## Annotation

see what is avaliable:

```{r}
columns(org.Hs.eg.db)
```

```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts), # Our genenames
                     keytype="ENSEMBL",        # The format of our genenames
                     column="SYMBOL",          # The new format we want to add
                     multiVals="first")
```

```{r}
head(mycounts)
```

# DESeq2 analysis:

```{r}
library(DESeq2)
```

set up our object for DESeq analysis
```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds
```

```{r}
dds
```


```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```


# Volcono plot

```{r}
plot(res$log2FoldChange,res$padj)
```

```{r}
plot(res$log2FoldChange,-log(res$padj))
```


```{r}
mycols<- rep("gray",length(res$padj))
mycols[abs(res$log2FoldChange<2)]="blue"
mycols[(res$log2FoldChange < 0.05) ]="gray"
plot(res$log2FoldChange,-log(res$padj),col=mycols)
```






